---
jupyter:
  jupytext:
    text_representation:
      extension: .Rmd
      format_name: rmarkdown
      format_version: '1.2'
      jupytext_version: 1.16.4
  kernelspec:
    display_name: R
    language: R
    name: ir
---

## Simple REST API

Make a request to documented address and get a response from same address.

Today, we work on the [UK Police Data API](https://data.police.uk/docs/).

```{r vscode={'languageId': 'r'}}
library("httr2")
```

```{r vscode={'languageId': 'r'}}
parameters <- list(
    lat = 52.629729,
    lng = -1.131592,
    date = "2024-01"
)

r = request("https://data.police.uk/api/crimes-street/all-crime")
r <- req_url_query(r, !!! parameters)
response = req_perform( r ) 

print( resp_body_json(response) )
```

<!-- #region -->
### Tasks

* Calculate how many crimes happened this year from the API above at London.
* At London, how many stop and searches took place?
* For crimes at Oxford, what outcomes there are?
* For crimes in London, identify the neighbourhood teams and print out for each their communication means - which communication mean is most common?


* Using [Cats as API](https://cataas.com/), print URLs for ten black cats and ten orange cats
<!-- #endregion -->



```{r vscode={'languageId': 'r'}}

```

# Simple API with authentication

We use [Abstract API](https://docs.abstractapi.com/holidays) to get holidays per country.
The API uses **API key** for authentication.

```{r vscode={'languageId': 'r'}}
params <- list(
  api_key = "THIS IS A SECRET",
  country = "FI",
  year = 2030,
  month = 12,
  day = 25
)

r <- request("https://holidays.abstractapi.com/v1/")
r <- req_url_query(r, !!! parameters)
response = req_perform( r ) 

print( resp_body_json(response) )
```

### Tasks

* How many holidays Finland has in July of current year?
* How many holidays United States has in July of current year?
* How many holidays Germany has in July of current year?

```{r vscode={'languageId': 'r'}}

```

### University of Helsinki buildings

Use the University of [Helsinki API gateway](https://api.helsinki.fi/portal/user/login) to create your own application, request access to the building API and then calculating how many buildings university has.

```{r vscode={'languageId': 'r'}}
header <- list(
  "X-Api-Key" = "THIS IS A SECRET"
)

r <- request("https://gw.api.helsinki.fi/building/v1/building/list")
r <- req_headers(r, !!! parameters)
response = req_perform( r )

print( data.text )
```

## OAuth

For OAuth on its simples form, we use the [OAuth Playground](https://www.oauth.com/playground/) to mimic a full OAuth flow.

1. Use Authorization Code -flow (for simplicity), you need to register a client.
1. Store client id, client secret, user login and user password somewhere.
1. Open the link provided, the username and password should be the same.

```{r vscode={'languageId': 'r'}}
## this authenticates you as the correct developer

client_id <- "your_client_id"
client_secret <- "your_client_secret"

```

```{r vscode={'languageId': 'r'}}
oauth_client <- oauth_client(
  id = client_id,
  secret = client_secret,
  token_url = "https://www.example.com/token",
)

oauth_flow_auth_code( oauth_client , auth_url = "https://www.oauth.com/playground/auth-dialog.html", redirect_uri="https://www.oauth.com/playground/authorization-code.html" )
```

### Task

* Use OAuth2 API from some real web service (such as X) and collect data.



